---
title: "GSS challange More Data"
author: "Evelyn Chairez"
date: "Summer 2024"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

This is a type of categorical problem, with at least one problem with multiple levels. To account for this we will use a Fisher Test

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
dd3 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
In this graph, the first slide displays the number of individuals categorized by races, white. black, and others. The white race had the most with over 1500 reports, and the black and other races had less than 250 reports. 
In the last display the chart shows the percentage of the races in favor of capital punishment. Each category was compared to each other, the white compared to support versus the opposition of capitol punishment and was in favor of capital punishment by approximately 64%. The black race was in opposition with a much lower percentage of around 57%. The other race was in favor of capital punishment with a near report for its opposite, the support percentage was around 55%. 

## Now with The Most data (fewer variables gives more data points)



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```
  
The summary displays that 65% of the white race was reported, 42% of the black race, and 55% of the other race were in favor were all in favor of capitol punishment. This is compared to 34% of the white race in opposition of capital punishment, 57% of the black race. and 44% of the other race. 


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table2)
chisqtestGC(table2)
```


## More data version
  
```{r}
fisherXY(mydata$race, mydata$cappun, printAll = "TRUE")
```
  
In the inferential data, we ran a CHIC square test and a Fisher test to get multiple results of each comparison. The first test, the CHIC square shared that the probability it pulled was equal to less than .000000000000000022. this number is too small. 
The CHI squared results align with the Fisher test and the comparable results show that the probability is too low to conclude against the hypothesis. We will reject the null hypothesis on these and will continue compiling data and information for a comparative analysis. 

  
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
    
To conclude this report we are looking at the compiled data and asking ourselves if it matches the hypothesis prediction. In this first area of data, we found that the number of people in the white race category overwhelmed the data for the other two races. This created a large margin of differing data and had a statistical significance. This numerical data summed it up into a two leveled comparison with each race being compared to the next. In this case, the Fisher Test and the CHI squared test verified that each result's probability was calculated. The results argued in favor of the hypothesis and showed how race had severe implications in determining if it had the significance of capital punishment. Referring to the second comparison on the second branch, we will conclude the following information in that format. 

